### 2013-05-04 trunk/Paper/main.tex
add paper docs

%% bare_jrnl_compsoc.tex
%% V1.3
%% 2007/01/11
%% by Michael Shell
%% See:
%% http://www.michaelshell.org/
%% for current contact information.
%%
%% This is a skeleton file demonstrating the use of IEEEtran.cls
%% (requires IEEEtran.cls version 1.7 or later) with an IEEE Computer
%% Society journal paper.
%%
%% Support sites:
%% http://www.michaelshell.org/tex/ieeetran/
%% http://www.ctan.org/tex-archive/macros/latex/contrib/IEEEtran/
%% and
%% http://www.ieee.org/

%%*************************************************************************
%% Legal Notice:
%% This code is offered as-is without any warranty either expressed or
%% implied; without even the implied warranty of MERCHANTABILITY or
%% FITNESS FOR A PARTICULAR PURPOSE! 
%% User assumes all risk.
%% In no event shall IEEE or any contributor to this code be liable for
%% any damages or losses, including, but not limited to, incidental,
%% consequential, or any other damages, resulting from the use or misuse
%% of any information contained here.
%%
%% All comments are the opinions of their respective authors and are not
%% necessarily endorsed by the IEEE.
%%
%% This work is distributed under the LaTeX Project Public License (LPPL)
%% ( http://www.latex-project.org/ ) version 1.3, and may be freely used,
%% distributed and modified. A copy of the LPPL, version 1.3, is included
%% in the base LaTeX documentation of all distributions of LaTeX released
%% 2003/12/01 or later.
%% Retain all contribution notices and credits.
%% ** Modified files should be clearly indicated as such, including  **
%% ** renaming them and changing author support contact information. **
%%
%% File list of work: IEEEtran.cls, IEEEtran_HOWTO.pdf, bare_adv.tex,
%%                    bare_conf.tex, bare_jrnl.tex, bare_jrnl_compsoc.tex
%%*************************************************************************

% *** Authors should verify (and, if needed, correct) their LaTeX system  ***
% *** with the testflow diagnostic prior to trusting their LaTeX platform ***
% *** with production work. IEEE's font choices can trigger bugs that do  ***
% *** not appear when using other class files.                            ***
% The testflow support page is at:
% http://www.michaelshell.org/tex/testflow/

% Note that the a4paper option is mainly intended so that authors in
% countries using A4 can easily print to A4 and see how their papers will
% look in print - the typesetting of the document will not typically be
% affected with changes in paper size (but the bottom and side margins will).
% Use the testflow package mentioned above to verify correct handling of
% both paper sizes by the user's LaTeX system.
%
% Also note that the "draftcls" or "draftclsnofoot", not "draft", option
% should be used if it is desired that the figures are to be displayed in
% draft mode.
%
% The Computer Society usually requires 12pt for submissions.
%
\documentclass[12pt,journal,compsoc]{IEEEtran}
%
% If IEEEtran.cls has not been installed into the LaTeX system files,
% manually specify the path to it like:
% \documentclass[12pt,journal,compsoc]{../sty/IEEEtran}

% Some very useful LaTeX packages include:
% (uncomment the ones you want to load)

% *** MISC UTILITY PACKAGES ***
%
%\usepackage{ifpdf}
% Heiko Oberdiek's ifpdf.sty is very useful if you need conditional
% compilation based on whether the output is pdf or dvi.
% usage:
% \ifpdf
%   % pdf code
% \else
%   % dvi code
% \fi
% The latest version of ifpdf.sty can be obtained from:
% http://www.ctan.org/tex-archive/macros/latex/contrib/oberdiek/
% Also, note that IEEEtran.cls V1.7 and later provides a builtin
% \ifCLASSINFOpdf conditional that works the same way.
% When switching from latex to pdflatex and vice-versa, the compiler may
% have to be run twice to clear warning/error messages.

% *** CITATION PACKAGES ***
%
\ifCLASSOPTIONcompsoc
  % IEEE Computer Society needs nocompress option
  % requires cite.sty v4.0 or later (November 2003)
  % \usepackage[nocompress]{cite}
\else
  % normal IEEE
  % \usepackage{cite}
\fi
% cite.sty was written by Donald Arseneau
% V1.6 and later of IEEEtran pre-defines the format of the cite.sty package
% \cite{} output to follow that of IEEE. Loading the cite package will
% result in citation numbers being automatically sorted and properly
% "compressed/ranged". e.g., [1], [9], [2], [7], [5], [6] without using
% cite.sty will become [1], [2], [5]--[7], [9] using cite.sty. cite.sty's
% \cite will automatically add leading space, if needed. Use cite.sty's
% noadjust option (cite.sty V3.8 and later) if you want to turn this off.
% cite.sty is already installed on most LaTeX systems. Be sure and use
% version 4.0 (2003-05-27) and later if using hyperref.sty. cite.sty does
% not currently provide for hyperlinked citations.
% The latest version can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/cite/
% The documentation is contained in the cite.sty file itself.
%
% Note that some packages require special options to format as the Computer
% Society requires. In particular, Computer Society  papers do not use
% compressed citation ranges as is done in typical IEEE papers
% (e.g., [1]-[4]). Instead, they list every citation separately in order
% (e.g., [1], [2], [3], [4]). To get the latter we need to load the cite
% package with the nocompress option which is supported by cite.sty v4.0
% and later. Note also the use of a CLASSOPTION conditional provided by
% IEEEtran.cls V1.7 and later.

% *** GRAPHICS RELATED PACKAGES ***
%
\ifCLASSINFOpdf
  % \usepackage[pdftex]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../pdf/}{../jpeg/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
\else
  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
  % will default to the driver specified in the system graphics.cfg if no
  % driver is specified.
  % \usepackage[dvips]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../eps/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.eps}
\fi
% graphicx was written by David Carlisle and Sebastian Rahtz. It is
% required if you want graphics, photos, etc. graphicx.sty is already
% installed on most LaTeX systems. The latest version and documentation can
% be obtained at: 
% http://www.ctan.org/tex-archive/macros/latex/required/graphics/
% Another good source of documentation is "Using Imported Graphics in
% LaTeX2e" by Keith Reckdahl which can be found as epslatex.ps or
% epslatex.pdf at: http://www.ctan.org/tex-archive/info/
%
% latex, and pdflatex in dvi mode, support graphics in encapsulated
% postscript (.eps) format. pdflatex in pdf mode supports graphics
% in .pdf, .jpeg, .png and .mps (metapost) formats. Users should ensure
% that all non-photo figures use a vector format (.eps, .pdf, .mps) and
% not a bitmapped formats (.jpeg, .png). IEEE frowns on bitmapped formats
% which can result in "jaggedy"/blurry rendering of lines and letters as
% well as large increases in file sizes.
%
% You can find documentation about the pdfTeX application at:
% http://www.tug.org/applications/pdftex

% *** MATH PACKAGES ***
%
%\usepackage[cmex10]{amsmath}
% A popular package from the American Mathematical Society that provides
% many useful and powerful commands for dealing with mathematics. If using
% it, be sure to load this package with the cmex10 option to ensure that
% only type 1 fonts will utilized at all point sizes. Without this option,
% it is possible that some math symbols, particularly those within
% footnotes, will be rendered in bitmap form which will result in a
% document that can not be IEEE Xplore compliant!
%
% Also, note that the amsmath package sets \interdisplaylinepenalty to 10000
% thus preventing page breaks from occurring within multiline equations. Use:
%\interdisplaylinepenalty=2500
% after loading amsmath to restore such page breaks as IEEEtran.cls normally
% does. amsmath.sty is already installed on most LaTeX systems. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/required/amslatex/math/

% *** SPECIALIZED LIST PACKAGES ***
%
%\usepackage{algorithmic}
% algorithmic.sty was written by Peter Williams and Rogerio Brito.
% This package provides an algorithmic environment fo describing algorithms.
% You can use the algorithmic environment in-text or within a figure
% environment to provide for a floating algorithm. Do NOT use the algorithm
% floating environment provided by algorithm.sty (by the same authors) or
% algorithm2e.sty (by Christophe Fiorio) as IEEE does not use dedicated
% algorithm float types and packages that provide these will not provide
% correct IEEE style captions. The latest version and documentation of
% algorithmic.sty can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/algorithms/
% There is also a support site at:
% http://algorithms.berlios.de/index.html
% Also of interest may be the (relatively newer and more customizable)
% algorithmicx.sty package by Szasz Janos:
% http://www.ctan.org/tex-archive/macros/latex/contrib/algorithmicx/

% *** ALIGNMENT PACKAGES ***
%
%\usepackage{array}
% Frank Mittelbach's and David Carlisle's array.sty patches and improves
% the standard LaTeX2e array and tabular environments to provide better
% appearance and additional user controls. As the default LaTeX2e table
% generation code is lacking to the point of almost being broken with
% respect to the quality of the end results, all users are strongly
% advised to use an enhanced (at the very least that provided by array.sty)
% set of table tools. array.sty is already installed on most systems. The
% latest version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/required/tools/

%\usepackage{mdwmath}
%\usepackage{mdwtab}
% Also highly recommended is Mark Wooding's extremely powerful MDW tools,
% especially mdwmath.sty and mdwtab.sty which are used to format equations
% and tables, respectively. The MDWtools set is already installed on most
% LaTeX systems. The lastest version and documentation is available at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/mdwtools/

% IEEEtran contains the IEEEeqnarray family of commands that can be used to
% generate multiline equations as well as matrices, tables, etc., of high
% quality.

%\usepackage{eqparbox}
% Also of notable interest is Scott Pakin's eqparbox package for creating
% (automatically sized) equal width boxes - aka "natural width parboxes".
% Available at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/eqparbox/

% *** SUBFIGURE PACKAGES ***
%\ifCLASSOPTIONcompsoc
%\usepackage[tight,normalsize,sf,SF]{subfigure}
%\else
%\usepackage[tight,footnotesize]{subfigure}
%\fi
% subfigure.sty was written by Steven Douglas Cochran. This package makes it
% easy to put subfigures in your figures. e.g., "Figure 1a and 1b". For IEEE
% work, it is a good idea to load it with the tight package option to reduce
% the amount of white space around the subfigures. Computer Society papers
% use a larger font and \sffamily font for their captions, hence the
% additional options needed under compsoc mode. subfigure.sty is already
% installed on most LaTeX systems. The latest version and documentation can
% be obtained at:
% http://www.ctan.org/tex-archive/obsolete/macros/latex/contrib/subfigure/
% subfigure.sty has been superceeded by subfig.sty.

%\ifCLASSOPTIONcompsoc
%  \usepackage[caption=false]{caption}
%  \usepackage[font=normalsize,labelfont=sf,textfont=sf]{subfig}
%\else
%  \usepackage[caption=false]{caption}
%  \usepackage[font=footnotesize]{subfig}
%\fi
% subfig.sty, also written by Steven Douglas Cochran, is the modern
% replacement for subfigure.sty. However, subfig.sty requires and
% automatically loads Axel Sommerfeldt's caption.sty which will override
% IEEEtran.cls handling of captions and this will result in nonIEEE style
% figure/table captions. To prevent this problem, be sure and preload
% caption.sty with its "caption=false" package option. This is will preserve
% IEEEtran.cls handing of captions. Version 1.3 (2005/06/28) and later 
% (recommended due to many improvements over 1.2) of subfig.sty supports
% the caption=false option directly:
%\ifCLASSOPTIONcompsoc
%  \usepackage[caption=false,font=normalsize,labelfont=sf,textfont=sf]{subfig}
%\else
%  \usepackage[caption=false,font=footnotesize]{subfig}
%\fi
%
% The latest version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/subfig/
% The latest version and documentation of caption.sty can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/caption/

% *** FLOAT PACKAGES ***
%
%\usepackage{fixltx2e}
% fixltx2e, the successor to the earlier fix2col.sty, was written by
% Frank Mittelbach and David Carlisle. This package corrects a few problems
% in the LaTeX2e kernel, the most notable of which is that in current
% LaTeX2e releases, the ordering of single and double column floats is not
% guaranteed to be preserved. Thus, an unpatched LaTeX2e can allow a
% single column figure to be placed prior to an earlier double column
% figure. The latest version and documentation can be found at:
% http://www.ctan.org/tex-archive/macros/latex/base/

%\usepackage{stfloats}
% stfloats.sty was written by Sigitas Tolusis. This package gives LaTeX2e
% the ability to do double column floats at the bottom of the page as well
% as the top. (e.g., "\begin{figure*}[!b]" is not normally possible in
% LaTeX2e). It also provides a command:
%\fnbelowfloat
% to enable the placement of footnotes below bottom floats (the standard
% LaTeX2e kernel puts them above bottom floats). This is an invasive package
% which rewrites many portions of the LaTeX2e float routines. It may not work
% with other packages that modify the LaTeX2e float routines. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/sttools/
% Documentation is contained in the stfloats.sty comments as well as in the
% presfull.pdf file. Do not use the stfloats baselinefloat ability as IEEE
% does not allow \baselineskip to stretch. Authors submitting work to the
% IEEE should note that IEEE rarely uses double column equations and
% that authors should try to avoid such use. Do not be tempted to use the
% cuted.sty or midfloat.sty packages (also by Sigitas Tolusis) as IEEE does
% not format its papers in such ways.

%\ifCLASSOPTIONcaptionsoff
%  \usepackage[nomarkers]{endfloat}
% \let\MYoriglatexcaption\caption
% \renewcommand{\caption}[2][\relax]{\MYoriglatexcaption[#2]{#2}}
%\fi
% endfloat.sty was written by James Darrell McCauley and Jeff Goldberg.
% This package may be useful when used in conjunction with IEEEtran.cls'
% captionsoff option. Some IEEE journals/societies require that submissions
% have lists of figures/tables at the end of the paper and that
% figures/tables without any captions are placed on a page by themselves at
% the end of the document. If needed, the draftcls IEEEtran class option or
% \CLASSINPUTbaselinestretch interface can be used to increase the line
% spacing as well. Be sure and use the nomarkers option of endfloat to
% prevent endfloat from "marking" where the figures would have been placed
% in the text. The two hack lines of code above are a slight modification of
% that suggested by in the endfloat docs (section 8.3.1) to ensure that
% the full captions always appear in the list of figures/tables - even if
% the user used the short optional argument of \caption[]{}.
% IEEE papers do not typically make use of \caption[]'s optional argument,
% so this should not be an issue. A similar trick can be used to disable
% captions of packages such as subfig.sty that lack options to turn off
% the subcaptions:
% For subfig.sty:
% \let\MYorigsubfloat\subfloat
% \renewcommand{\subfloat}[2][\relax]{\MYorigsubfloat[]{#2}}
% For subfigure.sty:
% \let\MYorigsubfigure\subfigure
% \renewcommand{\subfigure}[2][\relax]{\MYorigsubfigure[]{#2}}
% However, the above trick will not work if both optional arguments of
% the \subfloat/subfig command are used. Furthermore, there needs to be a
% description of each subfigure *somewhere* and endfloat does not add
% subfigure captions to its list of figures. Thus, the best approach is to
% avoid the use of subfigure captions (many IEEE journals avoid them anyway)
% and instead reference/explain all the subfigures within the main caption.
% The latest version of endfloat.sty and its documentation can obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/endfloat/
%
% The IEEEtran \ifCLASSOPTIONcaptionsoff conditional can also be used
% later in the document, say, to conditionally put the References on a 
% page by themselves.

% *** PDF, URL AND HYPERLINK PACKAGES ***
%
%\usepackage{url}
% url.sty was written by Donald Arseneau. It provides better support for
% handling and breaking URLs. url.sty is already installed on most LaTeX
% systems. The latest version can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/misc/
% Read the url.sty source comments for usage information. Basically,
% \url{my_url_here}.

% *** Do not adjust lengths that control margins, column widths, etc. ***
% *** Do not use packages that alter fonts (such as pslatex).         ***
% There should be no need to do such things with IEEEtran.cls V1.6 and later.
% (Unless specifically asked to do so by the journal or conference you plan
% to submit to, of course. )

% correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor}

\usepackage{graphicx}
\begin{document}
%
% paper title
% can use linebreaks \\ within to get better formatting as desired
\title{Modelo de Simulac\'{i}on de circulaci\'{o}n Veh\'{i}culos\\ en las calles de San Jos\'{e}, Costa Rica}
%
%
% author names and IEEE memberships
% note positions of commas and nonbreaking spaces ( ~ ) LaTeX will not break
% a structure at a ~ so this keeps an author's name from being broken across
% two lines.
% use \thanks{} to gain access to the first footnote area
% a separate \thanks must be used for each paragraph as LaTeX2e's \thanks
% was not built to handle multiple paragraphs
%
%
%\IEEEcompsocitemizethanks is a special \thanks that produces the bulleted
% lists the Computer Society journals use for "first footnote" author
% affiliations. Use \IEEEcompsocthanksitem which works much like \item
% for each affiliation group. When not in compsoc mode,
% \IEEEcompsocitemizethanks becomes like \thanks and
% \IEEEcompsocthanksitem becomes a line break with idention. This
% facilitates dual compilation, although admittedly the differences in the
% desired content of \author between the different types of papers makes a
% one-size-fits-all approach a daunting prospect. For instance, compsoc 
% journal papers have the author affiliations above the "Manuscript
% received ..."  text while in non-compsoc journals this is reversed. Sigh.

\author{Mag Yari~Guevara Rivera,
        Eddy Miguel~Ramirez Jimenez% <-this % stops a space
\IEEEcompsocitemizethanks{\IEEEcompsocthanksitem M. Shell is with the Department
of Electrical and Computer Engineering, Georgia Institute of Technology, Atlanta,
GA, 30332.\protect\\
% note need leading \protect in front of \\ to get a newline within \thanks as
% \\ is fragile and will error, could use \hfil\break instead.
E-mail: see http://www.michaelshell.org/contact.html
\IEEEcompsocthanksitem J. Doe and J. Doe are with Anonymous University.}% <-this % stops a space
\thanks{Manuscript received April 19, 2005; revised January 11, 2007.}}

% note the % following the last \IEEEmembership and also \thanks - 
% these prevent an unwanted space from occurring between the last author name
% and the end of the author line. i.e., if you had this:
% 
% \author{....lastname \thanks{...} \thanks{...} }
%                     ^------------^------------^----Do not want these spaces!
%
% a space would be appended to the last name and could cause every name on that
% line to be shifted left slightly. This is one of those "LaTeX things". For
% instance, "\textbf{A} \textbf{B}" will typeset as "A B" not "AB". To get
% "AB" then you have to do: "\textbf{A}\textbf{B}"
% \thanks is no different in this regard, so shield the last } of each \thanks
% that ends a line with a % and do not let a space in before the next \thanks.
% Spaces after \IEEEmembership other than the last one are OK (and needed) as
% you are supposed to have spaces between the names. For what it is worth,
% this is a minor point as most people would not even notice if the said evil
% space somehow managed to creep in.

% The paper headers
\markboth{Journal of \LaTeX\ Class Files,~Vol.~6, No.~1, January~2007}%
{Shell \MakeLowercase{\textit{et al.}}: Bare Demo of IEEEtran.cls for Computer Society Journals}
% The only time the second header will appear is for the odd numbered pages
% after the title page when using the twoside option.
% 
% *** Note that you probably will NOT want to include the author's ***
% *** name in the headers of peer review papers.                   ***
% You can use \ifCLASSOPTIONpeerreview for conditional compilation here if
% you desire.

% The publisher's ID mark at the bottom of the page is less important with
% Computer Society journal papers as those publications place the marks
% outside of the main text columns and, therefore, unlike regular IEEE
% journals, the available text space is not reduced by their presence.
% If you want to put a publisher's ID mark on the page you can do it like
% this:
%\IEEEpubid{0000--0000/00\$00.00~\copyright~2007 IEEE}
% or like this to get the Computer Society new two part style.
%\IEEEpubid{\makebox[\columnwidth]{\hfill 0000--0000/00/\$00.00~\copyright~2007 IEEE}%
%\hspace{\columnsep}\makebox[\columnwidth]{Published by the IEEE Computer Society\hfill}}
% Remember, if you use this you must call \IEEEpubidadjcol in the second
% column for its text to clear the IEEEpubid mark (Computer Society jorunal
% papers don't need this extra clearance.)

% use for special paper notices
%\IEEEspecialpapernotice{(Invited Paper)}

\setlength\parindent{24pt}

% for Computer Society papers, we must declare the abstract and index terms
% PRIOR to the title within the \IEEEcompsoctitleabstractindextext IEEEtran
% command as these need to go into the title area created by \maketitle.
\IEEEcompsoctitleabstractindextext{%
\begin{abstract}
%\boldmath
The abstract goes here.
\end{abstract}
% IEEEtran.cls defaults to using nonbold math in the Abstract.
% This preserves the distinction between vectors and scalars. However,
% if the journal you are submitting to favors bold math in the abstract,
% then you can use LaTeX's standard command \boldmath at the very start
% of the abstract to achieve this. Many IEEE journals frown on math
% in the abstract anyway. In particular, the Computer Society does
% not want either math or citations to appear in the abstract.

% Note that keywords are not normally used for peerreview papers.
\begin{IEEEkeywords}
Computer Society, ANN, Simulation, Traffic Lights, Model.
\end{IEEEkeywords}}

% make the title area
\maketitle

% To allow for easy dual compilation without having to reenter the
% abstract/keywords data, the \IEEEcompsoctitleabstractindextext text will
% not be used in maketitle, but will appear (i.e., to be "transported")
% here as \IEEEdisplaynotcompsoctitleabstractindextext when compsoc mode
% is not selected <OR> if conference mode is selected - because compsoc
% conference papers position the abstract like regular (non-compsoc)
% papers do!
\IEEEdisplaynotcompsoctitleabstractindextext
% \IEEEdisplaynotcompsoctitleabstractindextext has no effect when using
% compsoc under a non-conference mode.

% For peer review papers, you can put extra information on the cover
% page as needed:
% \ifCLASSOPTIONpeerreview
% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
% \fi
%
% For peerreview papers, this IEEEtran command inserts a page break and
% creates the second title. It will be ignored for other modes.
\IEEEpeerreviewmaketitle

\section{Introducci\'{o}n}

\section{Ambiente Presente en el Pa\'{i}s}
% Computer Society journal papers do something a tad strange with the very
% first section heading (almost always called "Introduction"). They place it
% ABOVE the main text! IEEEtran.cls currently does not do this for you.
% However, You can achieve this effect by making LaTeX jump through some
% hoops via something like:
%
%\ifCLASSOPTIONcompsoc
%  \noindent\raisebox{2\baselineskip}[0pt][0pt]%
%  {\parbox{\columnwidth}{\section{Introduction}\label{sec:introduction}%
%  \global\everypar=\everypar}}%
%  \vspace{-1\baselineskip}\vspace{-\parskip}\par
%\else
%  \section{Introduction}\label{sec:introduction}\par
%\fi
%
% Admittedly, this is a hack and may well be fragile, but seems to do the
% trick for me. Note the need to keep any \label that may be used right
% after \section in the above as the hack puts \section within a raised box.

% The very first letter is a 2 line initial drop letter followed
% by the rest of the first word in caps (small caps for compsoc).
% 
% form to use if the first word consists of a single letter:
% \IEEEPARstart{A}{demo} file is ....
% 
% form to use if you need the single drop letter followed by
% normal text (unknown if ever used by IEEE):
% \IEEEPARstart{A}{}demo file is ....
% 
% Some journals put the first two words in caps:
% \IEEEPARstart{T}{his demo} file is ....
% 
% Here we have the typical use of a "T" for an initial drop letter
% and "HIS" in caps to complete the first word.
\IEEEPARstart{E}{n} Costa Rica a lo largo de los a\~{n}os se ha podido observar como la gran
	cantidad de veh\'{i}culos automotores saturan sus principales ciudades, en especial
	el caso de su capital. Para el 2012 y seg\'{u}n el peri\'{o}dico La Naci\'{o}n
	alrededor de 19.000 buses y 260.000 autos luchan por un espacio para poder avanzar por las calles de San Jos\'{e}. \cite{Villegas2012}\\

Como parte de las medidas para mitigar problemas de embotellamientos, el
	Ministerio de Obras P\'{u}blicas y Transportes (MOPT) puso en funcionamiento,
	en el 2007, un sistema de sem\'{a}foros inteligentes el cual funcionaba para
	ese entonces en 180 intersecciones de la capital haciendo uso de unas 145
	c\'{a}maras aproximadamente. Sin embargo, en muchas ocasiones en las que se
	circul\'{o} por la capital, se pudo notar un problema en particular: en muchas
	situaciones un veh\'{i}culo no pod\'{i}a continuar avanzando, esto por el hecho
	de que el sem\'{a}foro de la siguiente intersecci\'{o}n estaba en rojo. A pesar
	de contar con dicho sistema inteligente se segu\'{i}an dando este tipo de
	problemas y para agravar la situaci\'{o}n, al pasar por estas calles no es
	posible encontrar alg\'{u}n tipo de sensor el cual le conceda la propiedad de inteligente a este sistema.\cite{Loaiza2007}

	Al persistir la presencia de embotellamientos, se presta m\'{a}s atenci\'{o}n a una serie situaciones presentadas.
	Por ejemplo, puede darse el escenario en el cual hay autom\'{o}viles esperando
	a que el sem\'{a}foro pase a verde, una vez ocurrido esto y al momento en que estos llegan a la
	siguiente intersecci\'{o}n el sem\'{a}foro de \'{e}sta sigue en rojo, esto
	podr\'{i}a hacer pensar que el sem\'{a}foro deber\'{i}a estar en verde justo antes de que lleguen, no obstante esto no ocurre siempre. Por otro lado en caso de estar esperando en una intersecci\'{o}n a que el sem\'{a}foro cambie, si no vienen autom\'{o}viles en el otro sentido, no deber\'{i}a mantenerse en luz roja y lo ideal ser\'{i}a que cambie a luz verde.

		Por \'{u}ltimo, cuando las luces de los sem\'{a}foros cambian en secuencia, aparentan
	saber que se aproximan carros hacia ellos y que por lo tanto deben realizar el
	cambio de luz, pero esto ocurre por coincidencias de los ciclos de luces
	asignados a cada sem\'{a}foro o son causa de una sincronizan llevada a cabo para realizar el cambio de luces de esta forma.

		Muchas de las interrogantes generadas a partir de los puntos anteriores,
		fueron contestadas luego de poder asistir a una exhibici\'{o}n del sistema de control de tr\'{a}nsito, localizado en las
	oficinas del Centro de Control de Tr\'{a}nsito del MOPT en San Jos\'{e}. Como resultado
	de esto qued\'{o} claro que no se trata de un sistema inteligente automatizado,
	sino de un sistema de control centralizado desde el cual los operadores del
	mismo pueden configurar los tiempos para los cambios de luces de los
	sem\'{a}foros dentro de la red.

% needed in second column of first page if using \IEEEpubid
%\IEEEpubidadjcol

% An example of a floating figure using the graphicx package.
% Note that \label must occur AFTER (or within) \caption.
% For figures, \caption should occur after the \includegraphics.
% Note that IEEEtran v1.7 and later has special internal code that
% is designed to preserve the operation of \label within \caption
% even when the captionsoff option is in effect. However, because
% of issues like this, it may be the safest practice to put all your
% \label just after \caption rather than within \caption{}.
%
% Reminder: the "draftcls" or "draftclsnofoot", not "draft", class
% option should be used if it is desired that the figures are to be
% displayed while in draft mode.
%
%\begin{figure}[!t]
%\centering
%\includegraphics[width=2.5in]{myfigure}
% where an .eps filename suffix will be assumed under latex, 
% and a .pdf suffix will be assumed for pdflatex; or what has been declared
% via \DeclareGraphicsExtensions.
%\caption{Simulation Results}
%\label{fig_sim}
%\end{figure}

% Note that IEEE typically puts floats only at the top, even when this
% results in a large percentage of a column being occupied by floats.
% However, the Computer Society has been known to put floats at the bottom.

% An example of a double column floating figure using two subfigures.
% (The subfig.sty package must be loaded for this to work.)
% The subfigure \label commands are set within each subfloat command, the
% \label for the overall figure must come after \caption.
% \hfil must be used as a separator to get equal spacing.
% The subfigure.sty package works much the same way, except \subfigure is
% used instead of \subfloat.
%
%\begin{figure*}[!t]
%\centerline{\subfloat[Case I]\includegraphics[width=2.5in]{subfigcase1}%
%\label{fig_first_case}}
%\hfil
%\subfloat[Case II]{\includegraphics[width=2.5in]{subfigcase2}%
%\label{fig_second_case}}}
%\caption{Simulation results}
%\label{fig_sim}
%\end{figure*}
%
% Note that often IEEE papers with subfigures do not employ subfigure
% captions (using the optional argument to \subfloat), but instead will
% reference/describe all of them (a), (b), etc., within the main caption.

% An example of a floating table. Note that, for IEEE style tables, the 
% \caption command should come BEFORE the table. Table text will default to
% \footnotesize as IEEE normally uses this smaller font for tables.
% The \label must come after \caption as always.
%
%\begin{table}[!t]
%% increase table row spacing, adjust to taste
%\renewcommand{\arraystretch}{1.3}
% if using array.sty, it might be a good idea to tweak the value of
% \extrarowheight as needed to properly center the text within the cells
%\caption{An Example of a Table}
%\label{table_example}
%\centering
%% Some packages, such as MDW tools, offer better commands for making tables
%% than the plain LaTeX2e tabular which is used here.
%\begin{tabular}{|c||c|}
%\hline
%One & Two\\
%\hline
%Three & Four\\
%\hline
%\end{tabular}
%\end{table}

% Note that IEEE does not put floats in the very first column - or typically
% anywhere on the first page for that matter. Also, in-text middle ("here")
% positioning is not used. Most IEEE journals use top floats exclusively.
% However, Computer Society journals sometimes do use bottom floats - bear
% this in mind when choosing appropriate optional arguments for the
% figure/table environments.
% Note that, LaTeX2e, unlike IEEE journals, places footnotes above bottom
% floats. This can be corrected via the \fnbelowfloat command of the
% stfloats package.

\section{Problem\'{a}tica}
	\subsection{El Problema y la soluci\'{o}n Actual}

		La gran cantidad de veh\'{i}culos automotores que circulan en las diferentes \'{a}reas
	de San Jos\'{e} provocan d\'{i}a a d\'{i}a embotellamientos que aumentan el tiempo requerido para ir de un punto a otro, as\'{i} como el consumo de combustible y por ende las emisiones de gases contaminantes.
	La situaci\'{o}n se presenta m\'{a}s alarmante cuando el mismo alcalde de San Jos\'{e}, el
	se\~{n}or Johnny Araya afirma que la cantidad de autos mencionados
	anteriormente ocupan un 70\% del espacio vial en la capital, pero \'{u}nicamente trasladan el 30\% del mill\'{o}n de personas que ingresan a San Jos\'{e} todos los d\'{i}as.\cite{Villegas2012}

		De acuerdo con la entrevista realizada a Iver Brade Monge, del Centro de
	Control de Tr\'{a}nsito del MOPT,  actualmente en Costa Rica el sistema con el que se cuenta es centralizado para permitir as\'{i} el control de los sem\'{a}foros, no obstante el proceso
	es manual y debe ser realizado por los operadores del mismo, los cuales de
	acuerdo con las estimaciones que realicen se aumenta o disminuye el tiempo de duraci\'{o}n de la luz verde del sem\'{a}foro, utilizando datos brindados por los contadores o c\'{a}maras localizados dentro de la red de sem\'{a}foros. Los datos de los contadores se emplean para determinar como aumenta o disminuye la cantidad de autom\'{o}viles que para por una determinada intersecci\'{o}n, mientras que las c\'{a}maras se emplean para poder visualizar la existencia o no de congestiones en las calles.

	\subsection{Infraestructura Vial}
	Quiz\'{a}s uno de los principales causantes o factores que propician los embotellamientos a nivel nacional, corresponde a la forma en que fueron dise\~{n}adas las calles sobre todo en la capital. Un recorrido por \'{e}stas muestra aspectos el ancho de las calles que por lo general poseen dos carriles, alguno de los cuales puede ser utilizado como zona de parqueo, de carga y descarga o simplemente para permitir que los autobuses, taxis o similares puedan dejar o llevar pasajeros. Present\'{a}ndose alguno de los casos mencionados, la capacidad de cada calle o avenida se reduce considerablemente y en algunos casos puede pasar de poseer dos carriles disponibles a s\'{o}lo uno.

	\subsection{Cultura y Sociedad en las V\'{i}as P\'{u}blicas}
		El no disponer de una flu\'{i}da circulaci\'{o}n dentro de estas \'{a}reas
		responde a diferentes motivos, unos son culturales tal como lo menciona el autor de la siguiente
	nota: \textit{''algunos choferes agravan los atascamientos debido a
	maniobras indebidas, al ignorar la luz roja de los sem\'{a}foros o cuando irrespetan las
	zonas prohibidas para estacionarse\ldots La gente no aplica la cortes\'{i}a, no
	tiene paciencia y todo eso va perjudicando.''}
	\cite{Villegas2012}

		A pesar de la existencia de legislaciones que buscan mejorar las condiciones dentro de las zonas m\'{a}s congestionadas de la capital, se han presentado en ciertas ocasiones en el que problemas en esta materia causando cambios por decisiones gubernamentales, tal y como ocurri\'{o} en junio del 2009 periodo durante el cual se dio la eliminaci\'{o}n, temporal, de la restricci\'{o}n vehicular para
	ingresar a San Jos\'{e} causando un aumento, para ese tiempo, del 25\% de
	veh\'{i}culos con personas que trataban de llegar a sus destinos dentro de la
	capital. Con eso no s\'{o}lo se dio incremento de automotores, sino que
	tambi\'{e}n se dieron aumentos en la duraci\'{o}n de las horas de mayor
	concentraci\'{o}n de los autom\'{o}viles y que de acuerdo con datos de
	ingenier\'{i}a de tr\'{a}nsito se estaban perdiendo entre un 10\% y 30\% en la
	disminuci\'{o}n del tiempo empleado por los autom\'{o}viles, as\'{i} como
	dejarse de obtener un ahorro de \$3 millones anuales en combustible.\cite{Mata2009}

		Si bien la restricci\'{o}n vehicular fue implementada como una forma de disminuir la cantidad de veh\'{i}culos automoteres, no todas las personas toman esto como un incentivo para hacer uso del transporte p\'{u}blico, si no que m\'{a}s bien lo ven como un obst\'{a}culo que los obliga a buscar una ruta alterna a la tradicional.

		No obstante, los eventos anteriores terminan vi\'{e}ndose intensificados  por la
	falta de una regulaci\'{o}n adecuada de los sem\'{a}foros. Aun contando con el sistema
	mencionado, no es seguro que se logre reducir los problemas para circular con
	fluidez dentro de los lugares m\'{a}s visitados, ya que la cantidad de veh\'{i}culos es
	cambiante, por lo cual al momento de realizar los ajustes mencionados, \'{e}sta puede estar variando de forma que se torna ineficiente la regulaci\'{o}n de dichos tiempos.

\section{El Modelo de Simulaci\'{o}n}
Como parte del trabajo a realizar se plantearon dos etapas importantes. La primera corresponde a un la realizaci\'{o}n del modelo de simulaci\'{o}n mencionado a continuaci\'{o}n, mientras que la segunda parte (no contemplada en el documento) presenta al combinaci\'{o}n de dicho modelo junto con la implemnetaci\'{o}n del modelo de red neuronal para que esta pueda ser evaluada mediante simulaciones de eventos.

\subsection{Detalle del Experimiento}
Dado la falta de un dato importante como lo son los tiempos de espera de los veh\'{i}culos, cuando circulan por las calles de San Jos\'{e}, resulta necesario llevar a cabo una simulaci\'{o}n de los cambios de luces para los sem\'{a}foros y del flujo de veh\'{i}culos presentes en el \'{a}rea.

Existen diferentes escenarios por los cuales se puedan dar los problemas planteados anteriormente, para este trabajo se utilizar\'{a} como zona de estudio la representada en la figura \ref{fig:traficoSJ}

	\begin{figure}[htp]
\centering
\includegraphics[scale=0.37]{images/trafico1.png}
\caption{Diagrama de Calles, San Jose Costa Rica (Google Maps)}
\label{fig:traficoSJ}
\end{figure}

		Basado en dicha imagen, se plantea el problema de cambio de las luces de los
	sem\'{a}foros tomando como referencia la \textbf{avenida:} 2, 4, 6 y 8,  y las
	\textbf{calles:}
	12, 10, 8 y 6. Para este caso la avenida segunda se ve afectada fuertemente por los veh\'{i}culos automotores que quedan en medio de las intersecciones, si bien este es un problema m\'{a}s cultural, se ver\'{i}a mitigado al contar con una red de sem\'{a}foros inteligentes que tomen decisiones basados en su entorno.

		As\'{i} por ejemplo, al presentarse una disminuci\'{o}n del flujo de veh\'{i}culos
	automotores desde la intersecci\'{o}n de la avenida 6 y calle 10 hasta la avenida
	segunda, debido a que los sem\'{a}foros de estas est\'{a}n pasando a la luz roja,
	posiblemente se generar\'{a} una obstrucci\'{o}n en los carros que circulan a  trav\'{e}s
	de la avenida segunda. Al presentarse este escenario, muchos de los veh\'{i}culos no podr\'{a}n pasar aun cuando tengan la luz verde permiti\'{e}ndoselos.

	Dependiendo de las condiciones en ese momento, ya sea un alto n\'{u}mero de
	veh\'{i}culos automotores en la cercan\'{i}a o la presencia de una fuerte lluvia que
	dificulte la conducci\'{o}n, es probable que se genere un efecto en cadena el cual
	culminar\'{a} afectando, con igual o mayor rapidez, otras \'{a}reas.

	Es aqu\'{i} donde la primera parte de este trabajo (la simulaci\'{o}n) entra en juego. Para este caso la simulaci\'{o}n s\'{o}lo comprender\'{a} escenarios normales dentro de la zona, si ningun evento o factor que altere la presencia o el flujo de veh\'{i}culos dentro de la misma. Otro aspecto importante a considerar es que el periodo a evaluarse corresponde a la franja horaria de las 4pm a las 6pm, dado que seg\'{u}n estimaciones realizadas en el MOPT, presenta una tendencia de fluja m\'{a}s constante en comparaci\'{o}n con las horas de la ma\~{n}ana donde se obtienen picos pronunciados en menor tiempo, correspondientes al traslado de personas a su lugar de trabajo.

	Siguiendo estas consideraciones, la simulaci\'{o}n a implementar tomar\'{a} en cuenta la llegada de veh\'{i}culos a la zona por medio de las calles o avenidas que conducen hacia la zona, es decir, calles cuya direcci\'{o}n o sentido llevan a los conductores hacia el \'{a}rea mencionada. De igual forma, se debe garantizar la circulaci\'{o}n de los veh\'{i}culos en las calles. Por lo anterior, en caso de presentarse una obstrucci\'{o}n en alg\'{u}n carril los carros deben ser capaces de realizar el cambio a un carril adyacente, teniendo en cuenta que cada calle tiene al menos dos carriles disponibles, salvo los casos como la avenida segunda la cual cuenta hasta con cinco, o las calles y avenidas que poseean zonas de carga y descarga que este puede reducirse a s\'{o}lo uno, no obstante este caso s\'{o}lo se presenta una vez dentro de la zona.

	Finalmente, los veh\'{i}culos han de circular dentro del \'{a}rea y en cada intersecci\'{o}n tomar\'{a}n una decisi\'{o}n si continuan en l\'{i}nea recta o si dan vuelta en la calle, \'{e}sto basado en el destino que cada automotor tenga. Para todos los casos a saber: cambios de carril, llegada de veh\'{i}culos y tendencias de flujos de los mismo, se han de emplear distribuciones probabil\'{i}sticas que permitan determinar cada uno de estos.

	Por otro lado y como etapa posterior, se ha pensado en la incorporaci\'{o}n a la simulaci\'{o}n de diferentes factores que alteren el flujo normal del tr\'{a}fico, \'{e}stos pueden darse de diferentes formas ya sea generando obstrucciones u obligando a reducciones de velocidades, sin importar el caso todos repercuten en cierta forma las calles por medio de los congestionamientos que generan.

\subsection{Factores a Considerarse}
Bas\'{a}ndose en eventos y condiciones que tienden a presentarse en la mayor\'{i}a de zonas del pa\'{i}s, se determinaron los factores a ser tomados en cuenta para la realizaci\'{o}n de este experimento total y a ser empleado en conjunto con las simulaciones. Cada factor consta de dos a tres niveles a ser tomados en cuenta, dichos niveles corresponden a la dificultad que el evento conlleva para los otros veh\'{i}culos dentro de la zona. Los factores se explican a continuaci\'{o}n.

\subsubsection{Accidentes de Tr\'{a}nsito}
Para la selecci\'{o}n de estos se tom\'{o} como base el mapa de distribuci\'{o}n espacial de accidentes para la zona obtenida del “Estudio de la distribuci\'{o}n espacial de accidentes de tr\'{a}nsito con v\'{i}ctimas en el cant\'{o}n de San Jos\'{e}” el cual fue proporcionado por el Consejo de Seguridad Vial. Para estos casos se utilizaron calles o avenidas de la zona que, de acuerdo con el estudio, presentan una mayor concentraci\'{o}n de accidentes.

\subsubsection{Zonas de Parqueos}
La idea de este  factor corresponde a un hecho caracter\'{i}stico como lo son paradas de autobuses, zonas de descarga y terminales en las calles. Si se analiza considerando que las calles posean un m\'{a}ximo de 2 carriles se pueden notar los siguientes comportamientos: si se poseen terminales de buses, se habla de un carril pr\'{a}cticamente dedicado a estas, mientras que para el otro caso se poseen bloqueos temporales de un carril en ciertos puntos por un tiempo determinado como resultado de que los buses realicen paradas para subir o bajar pasajeros.

	Cabe mencionar que para este caso s\'{o}lo se considerar\'{a}n casos de paradas permanentes, por el hecho de que realizar evaluaciones de otros resulta bastante difícil debido al grado de variación que se presenta, tal es el caso de el parqueo temporal de vehículos el cual dependerá de cada persona, según las habilidades y experiencia con las que cuenten para tardar más o menos tiempo en estacionar un vehículo.

\subsubsection{Eventos Meteorol\'{o}gicos}
Este factor var\'{i}a mucho pero igual se encuentra presente dentro de los factores que afectan y resulta necesario ser considerado. Para este caso, la presencia de lluvia en San Jos\'{e} ser\'{a} la principal circunstancia a evaluar tomando como punto de partida el hecho de que al presentarse, los conductores tienden a reducir la velocidad con el fin de evitar accidentes por las dificultades para manejar que se presentan con la calle mojada o la poca visibilidad en caso de una lluvia fuerte.

\subsection{Distribuciones a Emplear}

\section{Trabajo Futuro}
Posterior a la realizaci\'{o}n del modelo de simulaci\'{o}n, en el proceso pretende lograr un modelo de red neuronal artificial que pueda ser implementada dentro de este red de sem\'{a}foros y el cual permita que cada sem\'{a}foro se comunique con otros (hermanos) localizados dentro de un rango de \'{a}rea considerable de forma que cada uno pueda tener una idea lo que ocurre en su entorno y no \'{u}nicamente la informaci\'{o}n que puede ser recolectada por medio de los sensores en la intersecci\'{o}n administrada por cada sem\'{a}foro.

\subsection{Uso de Redes Neuronales}
La idea de usar redes neuronales, se basa en el hecho de que actualmente los sem\'{a}foros no son realmente inteligentes a como fueron promocionados en un principio. Dado que hacen uso de tiempos fijos, preprogramados, las tendencias cambias del flujo puede no siempre lograr satisfacerse de forma adecuada. Si bien los flujos de veh\'{i}culos siguien ciertas tendencias, de alzas o bajas, esto no se presenta siempre y en algunos momentos puede requerirse de ajustes inmediatos seg\'{u}n datos que se hayan analizado.

Dato que las redes neuronales se entrenan para aprender a generalizar resultados, lo que se busca es lograr que esta luego de un entrenamiento pueda conocer los momentos en los que de se debe aumentar o disminuir el tiempo de la luz verde, basado en su conocimiento actual tanto de la informaci\'{o}n recopilada por sensores como el insumo brindado por otros sem\'{a}foros cercanos.

\section{Conclusion}
The conclusion goes here.

% if have a single appendix:
%\appendix[Proof of the Zonklar Equations]
% or
%\appendix  % for no appendix heading
% do not use \section anymore after \appendix, only \section*
% is possibly needed

% use appendices with more than one appendix
% then use \section to start each appendix
% you must declare a \section before using any
% \subsection or using \label (\appendices by itself
% starts a section numbered zero.)
%

\appendices
\section{Proof of the First Zonklar Equation}
Appendix one text goes here.

% you can choose not to have a title for an appendix
% if you want by leaving the argument blank
\section{}
Appendix two text goes here.

% use section* for acknowledgement
\ifCLASSOPTIONcompsoc
  % The Computer Society usually uses the plural form
  \section*{Acknowledgments}
\else
  % regular IEEE prefers the singular form
  \section*{Acknowledgment}
\fi

The authors would like to thank...

% Can use something like this to put references on a page
% by themselves when using endfloat and the captionsoff option.
\ifCLASSOPTIONcaptionsoff
  \newpage
\fi

% trigger a \newpage just before the given reference
% number - used to balance the columns on the last page
% adjust value as needed - may need to be readjusted if
% the document is modified later
%\IEEEtriggeratref{8}
% The "triggered" command can be changed if desired:
%\IEEEtriggercmd{\enlargethispage{-5in}}

% references section

% can use a bibliography generated by BibTeX as a .bbl file
% BibTeX documentation can be easily obtained at:
% http://www.ctan.org/tex-archive/biblio/bibtex/contrib/doc/
% The IEEEtran BibTeX style support page is at:
% http://www.michaelshell.org/tex/ieeetran/bibtex/
%\bibliographystyle{IEEEtran}
% argument is your BibTeX string definitions and bibliography database(s)
%\bibliography{IEEEabrv,../bib/paper}
%
% <OR> manually copy in the resultant .bbl file
% set second argument of \begin to the number of references
% (used to reserve space for the reference number labels box)
\begin{thebibliography}{1}

\bibitem{IEEEhowto:kopka}
H.~Kopka and P.~W. Daly, \emph{A Guide to \LaTeX}, 3rd~ed.\hskip 1em plus
  0.5em minus 0.4em\relax Harlow, England: Addison-Wesley, 1999.

\end{thebibliography}

% biography section
% 
% If you have an EPS/PDF photo (graphicx package needed) extra braces are
% needed around the contents of the optional argument to biography to prevent
% the LaTeX parser from getting confused when it sees the complicated
% \includegraphics command within an optional argument. (You could create
% your own custom macro containing the \includegraphics command to make things
% simpler here.)
%\begin{biography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{mshell}}]{Michael Shell}
% or if you just want to reserve a space for a photo:

\begin{IEEEbiography}{Michael Shell}
Biography text here.
\end{IEEEbiography}

% if you will not have a photo at all:
\begin{IEEEbiographynophoto}{John Doe}
Biography text here.
\end{IEEEbiographynophoto}

% insert where needed to balance the two columns on the last page with
% biographies
%\newpage

\begin{IEEEbiographynophoto}{Jane Doe}
Biography text here.
\end{IEEEbiographynophoto}

% You can push biographies down or up by placing
% a \vfill before or after them. The appropriate
% use of \vfill depends on what kind of text is
% on the last page and whether or not the columns
% are being equalized.

%\vfill

% Can be used to pull up biographies so that the bottom of the last one
% is flush with the other column.
%\enlargethispage{-5in}

% that's all folks
\end{document}