### 2013-10-01 tags/1.1/nn-traffic/body/bibl.tex
This contains improves from previus version and adds the neural network on test mode (not yet integrated)

\chapter{Bibliograf\'{i}a}
	\label{chap:bib}

\begin{itemize}
  \item \textbf\textbf{[1]} Dobles R., (2011) Consecuencias de la contaminación del aire
  y de la atmósfera del sector energético y tendencias de las emisiones contaminantes.
  Recupeardo de
  \url{http://www.amcham.co.cr/newsletters/201102/airecontamina.pdf}
  \item \textbf{[2]} Coxworth, B. (Septiembre, 2010) Self-regulating traffic lights would
  improve vehicle flow. Gizmag. Recuperado de
  \url{http://www.gizmag.com/self-regulating-traffic-lights-improve-vehicle-flow/16396/}
  \item \textbf{[3]} Gilmore J., Elibiary, K. (1993) Traffic Management Applications of
  Neural Networks. Recuperado de
  \url{http://www.aaai.org/Papers/Workshops/1993/WS-93-04/WS93-04-012.pdf}
  \item \textbf{[4]} Pramod K, Dongun S., Chima A. (2008) Automatic Traffic Light Control
  System. Recuperado desde \url{http://bit.ly/JMcl5R}
  \item \textbf{[5]}Centro de Control de Tráfico MOPT (2011) Datos obtenidos del sistema.
  \item \textbf{[6]} Kriesel, D. (2005) A Brief Introduction to Neural Networks.
  Recuperado del sitio \url{http://www.dkriesel.com/en/science/neural_networks}
  \item \textbf{[7]} Hopfield, J (1962, Enero) Neural Networks and physical systems with
  emergent collective computational abilities. Pasadena, California
  \item Ayad M., M.S. Ahmad, M.Z.M. Yusoff, Hammad B. (s,f) Using Genetic
  Algorithm for Traffic Light Control System with a Pedestrian Crossing
  recuperado de: \url{http://paper.ijcsns.org/07_book/200902/20090212.pdf}
  \item \textbf{[9]} Rajendra Akerkar (2005) Introduction to Artificial Intelligence.
  Prentice-Hall OF India Private Limited, New Delhi. Pag 1-4
  \item \textbf{[10]} Greenman C., (1998) Choreographing the Dance of Traffic Lights. New
  York Times. Recuperado de http://www.nytimes.com/1998/09/17/technology/choreographing-the-dance-of-traffic-lights.html
  \item \textbf{[11]} Kumar G., (2007) Artificial Neural Network and Its Application. New
  Delhi. Págs 2, 4 recuperado desde
  http://www.iasri.res.in/ebook/EBADAT/5-Modeling%20and%20Forecasting%20Techniques%20in%20Agriculture/5-ANN_GKJHA_2007.pdf
  \item \textbf{[12]} Gallant S., (1993) Neural Network Learning and Expert Systems. MIT
  Press. Pags 133
  \item \textbf{[14]} Loaiza V., (Noviembre, 2007) Capital estrena semáforos inteligentes
  en 180 intersecciones. La Nación. Recuperado de
  \url{http://wvw.nacion.com/ln_ee/2007/noviembre/28/pais1332560.html}
  \item \textbf{[15]} Mata A., (Junio,2009) Acceso libre a San José provoca colapso vial.
  La Nación, Costa Rica. Recuperado de
  \url{http://wvw.nacion.com/ln_ee/2009/junio/16/pais1997637.html}
  \item \textbf{[16]} Organisation for Economic Co-operation and Development, European
  Conference of Ministers of Transport, OECD/ECMT Transport Research Centre
  (2006) Speed Management. Recuperado de \url{http://bit.ly/KZjVZ9}
  \item \textbf{[17]} Fallas H., (2007) Choferes ven más presas en capital. La Nación.
  Recuperado de \url{http://wvw.nacion.com/ln_ee/2007/enero/29/pais976503.html}
  \item \textbf{[18]} Jameson et al Longo, Kasper. Harrison's Principles of internal
  medicine. McGraw Hill, 2011. URL
  \url{http://www.amazon.com/Harrisons-Principles-Internal-Medicine-Volumes/dp/007174889X/ref=sr_1_1?ie=UTF8&qid=1329942776&sr=8-1}
  \item \textbf{[19]} Hilera y Martínez. (1995) Redes Neuronales Artificiales.
  Fundamentos, Modelos y Aplicaciones. Addison Wesley
  \item \textbf{[20]} Graupe D., (2007) Principles of Artificial Neural Networks. World
  Scientific. Singapur
  \item \textbf{[21]} Dayan p., (s,f)  Unsupervised Learning. MIT. URL:
  \url{http://www.gatsby.ucl.ac.uk/~dayan/papers/dun99b.pdf}
  \item \textbf{[22]} Basogain X., (s,f) Redes Neuronales Artificiales y sus
  Aplicaciones. España. Url:
  \url{http://ocw.ehu.es/ensenanzas-tecnicas/redes-neuronales-artificiales-y-sus-aplicaciones/contenidos/pdf/libro-del-curso}
  \item \textbf{[23]} Srinivasan D., Chee M. et Long R., (2006) Neural Networks for
  RealTime Traffic Signal Control. IEEE. Singapur   Url:
  \url{http://www.jhuapl.edu/spsa/PDF-SPSA/Srinivasan_etal_IEEETITS06.pdf}
  \item \textbf{[24]} Ramírez E., (2012) Reconocimiento de patrones en juegos utilizando
  redes neuronales. Tecnológico de Costa Rica.
  \item \textbf{[25]} Fritsch. Modular neural networks for speech recognition, 1996. URL
  \url{http://www.
dtic.mil/cgi-bin/GetTRDoc?AD=ADA326090&Location=U2&doc=GetTRDoc.pdf}

  \item \textbf{[26]} Brinckerhoff P., Farradyne T., (2010, Abril) Evaluation of Active
  Traffic Management Strategies. Recuperado de
  \url{http://www.metrocouncil.org/planning/transportation/MHSIS/AppEATMmodel.pdf}
  \item \textbf{[27]} Ezeel S., (2010, Enero)  Explaning International IT Application
  Leadership: Intelligent Transportation Systems. The Information Technology \& Innovation Foundation. Url:
  \url{http://www.itif.org/files/2010-1-27-ITS_Leadership.pdf}
  \item \textbf{[28]} Turnbull K, (s,f) Transit benefits from advanced transportation
  management centers. Texas Transportation Institute, USA URL:
  \url{http://www.etcproceedings.org/paper/transit-benefits-from-advanced-transportation-management-centres-45}
  \item \textbf{[29]} \url{http://ntl.bts.gov/lib/jpodocs/edldocs1/13480/ch3.pdf}

\end{itemize}